Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-04-03 src/tex/ms.tex
Adding more comments

%\definecolor{rb4}{HTML}{27408B}
%\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
%\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
\newcommand{\ripple}{\texttt{ripple}}
\newcommand{\python}{\texttt{python}}
\newcommand{\jax}{\texttt{jax}}
\newcommand{\zdethp}{\texttt{zdethp}}
\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}

@@

% Main body with filler text
\section{Introduction} \label{sec:intro}

@@

In order to recalibrate the model, we need to define an objective function that
quantify the performance of the model. A common choice for such metric is the
\textit{mismatch} function \citep{husa2016frequency}. It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveform we are comparing, and $t_0$ and $\phi_0$
are time shift and phase shift respectively. The $\langle h_1, h_2 \rangle$ is
commonly referred as the inner product, which is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density(PSD) of noise from the instrument, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
the relevant maximum and minimum frequencies for the integration. The mismatch
is a quantity that is closely related to the mean square error (MSE) between the
two waveforms, as it is often considered a weighted version of the MSE.

\kw{Remind me why don't we use the hybrid waveforms again? Cause we couldn't do it?}
In the original calibration \citep{khan2016frequency,husa2016frequency},
training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
waveforms while the rest of the waveforms are taken from NR simulations.
Instead, we solely use NR waveforms for comparison, since most NR waveforms used
have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
which they are long enough to contain part of the inspiral segment and all
merger and ringdown frequency information. We take the frequency limits as
$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,
where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
the IMRPhenomD's frequency range, except the minimum frequency is set higher
than that in the original calibration due to NR length. When compared with
IMRPhenomC, the frequency range is slightly extended to have a higher maximum
frequency.

To compute the integral shown in Eq. \ref{eq:inner_prod}, one need to choose a
relevant noise PSD. Since a common goal for creating waveform approximent is to
deploy the waveform model in downstream data analysis tasks, a reasonable choice
for the noise PSD should be closely related to the instrument of interest. We
choose the zero-detuned high-power (\zdethp) noise PSD \kw{add citation} as our
first choice of noise PSD. Because the total mass of the system simple scales
the frequency of the waveform, we need to choose a corresponding mass scale to
match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
all waveforms, since this mass scale is common in LIGO observations. In addition
to a detector-dependent PSD, we are also interested in a more
instrument-agnostic case such that we can evaluate the effectiveness of the
recalibration without considering detector effect. In this case, we set the
noise PSD to be a constant value.

@@

To evaluate the effectiveness of our optimization approach, we analyze
approximately 530 NR waveforms from the SXS catalog in this study. We
specifically select waveforms with negligible eccentricity
(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
consistent with the constraints of the waveform model. As shown in
Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
waveforms fall within the parameter space defined by the training waveforms.
This suggests that these test waveforms can be used for direct comparison with
the original model.

We begin by investigating the impact of joint optimization on an individual
waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
outperforms the original waveform, particularly in the inspiral region, where
the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
selected one of the testing waveforms from the catalog presented in
\citep{khan2016frequency}.
% In general, most waveforms within the neighborhood of the training waveforms
% show improvement in mismatches.

We use a constant noise spectrum, $\mathcal{L}{\mathrm{mean}}$, to evaluate the
mismatch of all testing waveforms, and present the resulting distribution in
Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
in the median. When using $\mathcal{L}{\mathrm{fl}}$, we observe a similar
improvement with a 22.9\% decrease in the median. However, the distribution lacks
a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
Despite this, both distributions exhibit significant improvement, and the
optimization scheme remains unaffected.

% \kw{We then compute the mismatch for all testing waveforms and plot the
% distribution of the mismatch in figure X. We find the distribution...}

>>>>

\definecolor{rr}{RGB}{173, 37, 37}
%\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
%\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
\newcommand{\te}[1]{{\color{rr}[TE: #1 ]}}
\newcommand{\ripple}{\texttt{ripple}}
\newcommand{\python}{\texttt{python}}
\newcommand{\jax}{\texttt{jax}}
\newcommand{\zdethp}{\texttt{zdethp}}

@@

% Main body with filler text
\section{Introduction} \label{sec:intro}
\te{General points:
\begin{itemize}
	\item Overall the text needs to be refined
	\item Plot styles should be made more uniform
\end{itemize}
}

@@

To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.

In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms. The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations. Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information. We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length. When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency. \te{What is the frequency spacing?} \te{Why do we not just go to the cutoff of 0.2, or do the NR waveforms end earlier? Think these choices need more motivation.}

\te{I think this discussion needs to be much more precise. Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper. So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD. This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned high-power (\zdethp) noise spectrum and a constant spectrum. Since in any data analysis tasks, model-generated waveforms will ultimately be used to compare with GW observations, they are subjected to the sensitivity of detectors. Hence, it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than a non-weighted spectrum. However, with sensitivity curve involved, we have to choose a corresponding mass scale. Here, we simply choose $M=50M_{\odot}$ for all waveforms, since this mass scale is common in LIGO observations. 
\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}
On the other hand, we also recalibrated coefficients with the constant spectrum for a more general purpose.

@@

First, to judge the performance of the optimization scheme, we take $\sim530$ NR waveforms from the SXS catalog. We choose testing waveforms with negligible eccentricity (${e<2\times10^{-3}}$) and precession (${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model. We see that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters that are within the parameter space spanned by training waveforms. Hence, we can take these testing waveforms to compare with the original model.

We first examine the effect of joint optimization on a single waveform. In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error. We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison. 
% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.

Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end with a lower mismatch. Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease. Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement. We see that the median of distribution has a $22.9\%$ decrease. However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}. This is due to the problem mentioned in section \ref{subsec:optimization}. Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.

% \kw{We then compute the mismatch for all testing waveforms and plot the distribution of the mismatch in figure X. We find the distribution...}
